Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-06-05 ThesisJP/Background/background.tex
Final?

\subsection{GPUs in medical ultrasound imaging}
Ultrasound scanners generate an incredible amount of data. A modern system capable of 3D imaging will generate several tera-bytes of data inside the probe handle during a clinical scan. Fortunately the data rate is reduced to the level of 2D imaging before the data reaches the scanner. The data rate of a 2D scan is "only" in order of giga-bytes per second. To handle this amount of data, ultrasound imaging systems have, for the last twenty years, relied heavily on application-specific integrated circuits (\nom{ASIC}{Application-specific integrated circuit}) for all steps in the processing pipeline \cite{Thomenius}. However, as the computational power of generic processors have increased, the expensive ASICs have gradually been replaced by software \cite{Guracar2013}. One of  the last remaining ASICs in ultrasound scanners is used for beamforming \cite{Thomenius}. As seen from (\ref{eq:das_background}), beamforming involves a reduction of the data rate with a factor equal to the number of elements in the ultrasound probe. The input data rate is therefore very high. Nevertheless, real-time beamforming has recently been achieved on a GPU \cite{Song2012}, and a French company (SuperSonic Imagine) are utilizing GPUs in their plane wave beamforming algorithm \cite{Tanter2014}. This is the same technique as reported implemented on a GPU by Yiu \textit{et al.} \cite{Yiu2011}. We can therefore conclude that beamforming is finally moving into software too. The largest players in the medical ultrasound business, General Electric and Philips, are also well aware of where the trend is going \cite{Thomenius2012} \cite{Metz2011}.

@@

\subsection{Narrowband or broadband Capon beamforming}
Both the LCA and BS-Capon beamformers are based upon a detailed investigation of how Capon beamfoming behaves when applied to medical ultrasound images. For the LCA beamformer, the windows selected by the Capon beamformer were investigated and a special set of predefined windows were defined. If one looks closer at the windows selected for the LCA beamformer we see that half the set performs micro-steering of the mainlobe and the other half provides different mainlobe widths. The results from applying the LCA beamformer \cite[Fig.\,9]{Synnevag2011} and the success of the BS-Capon beamformer shows that the main benefit of applying Capon beamforming to medical ultrasound imaging is micro-steering of the mainlobe, and not attenuation of far off-axis backscatter.

The Capon beamformer is derived for narrowband applications and has traditionally been applied to medical ultrasound imaging "as is". Medical ultrasound imaging is, however, a broadband imaging modality and one could think that a broadband Capon beamformer would therefore be needed \cite{Holfort2009}. The investigation performed in \cite{Nilsen2009} showed, that with pre-delayed data, energy will be centered on-axis and as little as three near-axis beams in beamspace is therefore adequate in order to achieve similar performance for BS-Capon and Capon beamforming. The two-way beam will be wider for unfocused (plane wave) transmit beams \cite{Holfort2009, Austeng2011}, but steering and focusing on receive will still make the main amount of energy arrive on axis. As shown in \textbf{Paper III} it is also possible to micro-steer with phase delays (narrowband) instead of time delays (broadband) without introducing large errors. Since Capon beamforming mainly performs main-lobe shifts smaller than the beam width \cite{Synnevag2011, Nilsen2009}, the difference between a broadband and narrowband Capon beamformer will be minimal \cite{Diamantis2014}. Broadband Capon beamforming is also significantly more computationally demanding than narrowband Capon beamforming.

>>>>

\subsection{GPUs in medical ultrasound imaging}
Ultrasound scanners generate an incredible amount of data. A modern system capable of 3D imaging will generate several tera-bytes of data inside the probe handle during a clinical scan. Fortunately the data rate is reduced to the level of 2D imaging before the data reaches the scanner. The data rate of a 2D scan is "only" in order of giga-bytes per second. To handle this amount of data, ultrasound imaging systems have, for the last twenty years, relied heavily on application-specific integrated circuits (\nom{ASIC}{Application-specific integrated circuit}) for all steps in the processing pipeline \cite{Thomenius}. However, as the computational power of generic processors have increased, the expensive ASICs have gradually been replaced by software \cite{Guracar2013}. One of  the last remaining ASICs in ultrasound scanners is used for beamforming \cite{Thomenius}. As seen from (\ref{eq:das_background}), beamforming involves a reduction of the data rate with a factor equal to the number of elements in the ultrasound probe. The input data rate is therefore very high. Nevertheless, real-time beamforming has recently been achieved on a GPU \cite{Song2012}, and a French company (SuperSonic Imagine) are utilizing GPUs in their plane wave beamforming algorithm \cite{Tanter2014}. This is the same technique as reported implemented on a GPU by Yiu \textit{et al.} \cite{Yiu2011}. Software beamforming is also available in the Vantage research ultrasound system by Verasonics\footnote{http://www.verasonics.com/25}. We can therefore conclude that beamforming is finally moving into software too. The largest players in the medical ultrasound business, General Electric and Philips, are also well aware of where the trend is going \cite{Thomenius2012} \cite{Metz2011}.

@@

\subsection{Capon beamforming of focused broadband beams}
Both the LCA and BS-Capon beamformers are based upon a detailed investigation of how Capon beamfoming behaves when applied to medical ultrasound images. For the LCA beamformer, the windows selected by the Capon beamformer were investigated and a special set of predefined windows were defined. If one looks closer at the windows selected for the LCA beamformer we see that half the set performs micro-steering of the mainlobe and the other half provides different mainlobe widths. The results from applying the LCA beamformer \cite[Fig.\,9]{Synnevag2011} and the success of the BS-Capon beamformer shows that the main benefit of applying Capon beamforming to medical ultrasound imaging is micro-steering of the mainlobe, and not attenuation of far off-axis backscatter. This is mainly due to the focused beams used in medical ultrasound which highly limits the amount of energy arriving off axis. This focusing of energy also explain why Capon beamforming works on broadband ultrasound data.

The Capon beamformer is derived for narrowband applications and has traditionally been applied to medical ultrasound imaging "as is". Medical ultrasound imaging is, however, a broadband imaging modality and one could think that a broadband Capon beamformer would therefore be needed \cite{Holfort2009}. The investigation performed for the BS-Capon beamformer \cite{Nilsen2009} showed, that with pre-delayed data, energy will be centered on-axis and as little as three near-axis beams in beamspace is therefore adequate in order to achieve similar performance for BS-Capon and Capon beamforming. The two-way beam will be wider for unfocused (plane wave) transmit beams \cite{Holfort2009, Austeng2011}, but steering and focusing on receive will still make the main amount of energy arrive on axis. As shown in \textbf{Paper III} it is also possible to micro-steer with phase delays (narrowband) instead of time delays (broadband) without introducing large errors. Since Capon beamforming mainly performs main-lobe shifts smaller than the beam width \cite{Synnevag2011, Nilsen2009}, the difference between a broadband and narrowband Capon beamformer will be minimal \cite{Diamantis2014}. Broadband Capon beamforming is also significantly more computationally demanding than narrowband Capon beamforming.
